A Hitchhiker's Guide to Reproducible Research in R

### 2021-06-29 manuscript.Rmd
fix small mistakes

These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible.
An easy to use standard, on the other hand, is more likely to be employed across many research projects.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

@@

One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
Two major strategies are helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.

@@

Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format. Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.

@@

Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:

@@

This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.

@@

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].

>>>>

These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
These open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible.
An easy to use standard, on the other hand, is more likely to be employed across many research projects.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

@@

One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
Two major strategies are helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an sufficently describe how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.

@@

Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format. Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.

@@

Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
R Markdown was extended to hundreds of document formats which are often highly configurable by their metadata.
The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:

@@

This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all R Markdown  (all files with the ending `.Rmd`) and there respective repro metadata.
Every time you change the `repro`-metadata, change the output format, or add a new R Markdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.

@@

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019:
